# Missing values

We use a single file "monthly_flask_co2" as an example to analyze the missing data in our project and we can clearly see that there are a lot of missing values in this dataset.
```{r}
library(ggplot2)
library(tidyverse)
library(patchwork)
monthly_co2 = read.csv("./data/monthly_flask_co2.csv")
monthly_co2[1:10,]
```

Firstly, we analysis its missing patterns.
```{r}
plot_missing <- function(sets, percent){
  mysets <- sets
  ############## main plot heatmap ############## 
  missing_patterns <- data.frame(is.na(mysets)) %>%
    group_by_all() %>%
    count(name = "count", sort = TRUE) %>%
    ungroup()
  
  patterns_cases <- missing_patterns %>% select(-count) %>% 
    rownames_to_column("id") %>% gather(key, value, -id)

  #index of complete case pattern
  y_complete <- which(rowSums(select(missing_patterns, -count)) == 0)

  patterns_cases <- patterns_cases %>% mutate(missing = ifelse(id == y_complete, "none", ifelse(value == 1, "yes", "no")))
  
  #count of missing values for each column in the dataset
  missing_columns <- colSums(is.na(mysets)) %>%
      sort(decreasing = TRUE) %>% 
      data.frame()
  names(missing_columns) <- "n"
  missing_columns <- rownames_to_column(missing_columns, "key")

  x_text <- ceiling(ncol(mysets)/2)
  if (nrow(missing_patterns) > 1) {
    y_text <- nrow(missing_patterns) - y_complete + 1
  } else {
    y_text <- 1
  }
  
  patterns_cases$id <- factor(as.numeric(patterns_cases$id))
  patterns_cases <- merge(patterns_cases, missing_columns, by = "key", all = TRUE)

  g1 <- ggplot(patterns_cases, aes(x = fct_reorder(key, -n), y = fct_rev(id), fill = missing)) +
    geom_tile(color = "white") + 
    xlab("variable") +
    ylab("missing pattern") +
    scale_fill_manual(values = c("grey", "darkgrey", "purple")) +
    annotate("text", x = x_text, y = y_text, label = "Complete Cases") +
    theme(legend.position = "none")
    #theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  
  ############## missing pattern barchart ############## 
  if (percent) {
    missing_columns$n <- missing_columns$n /nrow(mysets) * 100
    missing_patterns$count <- missing_patterns$count /nrow(mysets) * 100
    text2 <- "% rows\n missing:"
    text3 <- "% rows"
    ylim2 <- c(0, 100)
  } else {
    text2 <- "num rows\n missing:"
    text3 <- "row count"
    ylim2 <- c(0, NA)
  }

  g2 <- ggplot(missing_columns, aes(fct_reorder(key, -n), n)) +
    geom_bar(stat="identity",fill="cornflowerblue") +
    xlab("") +
    ylab(text2) +
    ylim(ylim2) +
    ggtitle("Missing value patterns") +
    theme_bw() +
    theme(panel.grid.major.x = element_blank())
    #theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  
  ############## missing pattern count barchart ############## 
  missing_patterns <- missing_patterns %>% rownames_to_column("id") %>% mutate(complete = ifelse(id == y_complete, "yes", "no"))
  missing_patterns$id <- factor(as.numeric(missing_patterns$id))
  
  g3 <- ggplot(missing_patterns, aes(fct_rev(id), count, fill = complete)) +
    geom_bar(stat="identity") +
    xlab("") +
    ylab(text3) +
    ylim(ylim2) +
    scale_fill_manual(values = c("cornflowerblue", "blue")) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank(), legend.position = "none") +
    coord_flip() 
  
  ############## final plot ############## 
  g_final <- g2 + plot_spacer() + g1 + g3 +
    plot_layout(ncol = 2, nrow = 2, heights = c(1, 3), widths = c(3.5, 1))
  
  return(g_final)
}
```

```{r}
plot_missing(monthly_co2, TRUE)
plot_missing(monthly_co2, FALSE)
```

As shown in the above graph, each inspection station has NA values in its column, while year and month columns don't have missing data. The Alert station has the highest missing percentage and missing count while the Mauna Loa Observatory has the lowest missing percentage and missing count. What's more, the type of missing data is Missing at Random(MAR). We can prove it by using the data of Kermadec Island Station and New Zealand Station.

```{r}
monthly_co2 %>% select('Yr', 'ker') %>% 
  group_by(Yr) %>% mutate(count = sum(is.na(ker))) %>% 
  ungroup() %>% select('Yr', 'count') %>% unique() %>% 
  ggplot(aes(x = Yr, y = count)) +
  geom_bar(stat = 'identity') +
  ggtitle('Yearly missing data of ker')

monthly_co2 %>% select('Yr', 'nzd') %>% 
  group_by(Yr) %>% mutate(count = sum(is.na(nzd))) %>% 
  ungroup() %>% select('Yr', 'count') %>% unique() %>% 
  ggplot(aes(x = Yr, y = count)) +
  geom_bar(stat = 'identity') +
  ggtitle('Yearly missing data of nzd')

```

We can observe that before 1975, both stations don't have any data. This trend exists in almost every station. For most stations, the former data has a higher percentage of missing data. However, the missing patterns between stations are still different except for the former year. In Kermadec Island Station, the missing data around 2020 is more, while the missing data of New Zealand Station around 2020 shows little difference with other years. Thus, the missing data of our dataset is Missing At Random. Since each station starts and ends recording data at different times, that's why we don’t have many complete cases, but we can still observe the trend of co2 generated by each station over time according to its own timeline.
