# Results

## CO2 changes by time

```{r}
library(ggplot2)

yearly_co2_tidy = read.csv("./data/yearly_co2_tidy.csv")
ggplot(data = na.omit(yearly_co2_tidy), aes(x = Yr, y = CO2_concentration, color = station)) +
  geom_line() +
  ggtitle("CO2 concentrations from 1960 to 2021") +
  labs(x = "year", y = "CO2 concentration") + 
  theme_grey(16)
```
```{r}

```


From this graph, we can conclude that there are increasingly amount of C02 concentrations for all the stations on the whole year by year over time. And this trend is consistent for all sampling stations.


## Factors affecting the CO2 changes

### locations: different sampling stations

```{r}
library(tidyverse)
monthly_co2 <- read.csv("./data/monthly_flask_co2.csv")
monthly_co2_tidy <- monthly_co2 %>%
  pivot_longer(cols = -c(Yr,Mn), names_to = "station", 
               values_to = "CO2_concentration")
ggplot(monthly_co2_tidy, aes(x=fct_reorder(station,CO2_concentration,na.rm=TRUE), y=CO2_concentration)) +
  geom_boxplot() +
  coord_flip() +
  ggtitle("Monthly CO2 concentrations for each station") +
  labs(x = "station", y = "monthly CO2 concentration") + 
  theme_grey(16)
```

From this graph, we can conclude that the median monthly CO2 concentration values are similar accross all the stations except for the Alert Station. The Alert Station has a median of more than 400, while the median of others are approximately between 350 and 375. That's because the Alert Station started to record data only from 2011, so it has a small range. The highest monthly CO2 concentration also comes form the Alert Station, which is almost 425. Besides the Alert Station, Cape Kumukahi Station and American Samoa Station have higher monthly CO2 concentration than Kermadec Island Station than the other three stations. The data for Mauna Loa Observatory is the most sparse because it is the earliest station to record data, and that's why its median is the lowest. There are no outliers.

### Sea water features

The sea water is continuously changing as well. In the data, one feature of sea water is the salinity. 

```{r}
sw_berm <- read_csv("./data/BERM.csv")
sw_bats <- read_csv("./data/BATS.csv")
sw_hawi <- read_csv("./data/HAWI.csv")
sw_total <- rbind(sw_berm, sw_bats, sw_hawi)
ggplot(sw_total, aes(x=DIC,y=Salinity)) + ggtitle("Salinity v.s DIC in BERM") + facet_wrap(~Station) +
  geom_hex(bins=30) + scale_fill_gradient(low = "lightblue", high = "blue")
```

From the above graphs, we can see that there is somehow some relation between Salinity and carbon in sea water. Sea water with higher salinity tend to have larger amount of dissolved inorganic carbon (DIC). This is a potential factor to the CO2 concentration differences.

## CO2 changes during Covid-19 pandemic

```{r}
df_covid <- na.omit(yearly_co2_tidy)[260:285,]
ggplot(data = df_covid, aes(x = Yr, y = CO2_concentration, color = station)) +
  geom_line() +
  annotate("rect", xmin = 2020.2, xmax = 2021, alpha = .2, ymin = 402, ymax = 422) +
  labs(x = "year", y = "CO2 concentration") + 
  theme_grey(16)
```

The shaded area is the time period during the Covid-19 pandemic. We select the time interval starting from 2020 March to present. And we want to focus on this interval to see if the pandemic affects the CO2 emission.
